Red Wines Exploration by Hadrien Lacroix
========================================================

References:
  P. Cortez, A. Cerdeira, F. Almeida, T. Matos and J. Reis. 
  Modeling wine preferences by data mining from physicochemical properties.
  In Decision Support Systems, Elsevier, 47(4):547-553. ISSN: 0167-9236.
  
More information about this dataset can be found [here](https://s3.amazonaws.com/udacity-hosted-downloads/ud651/wineQualityInfo.txt).

This tidy dataset contains 1,599 red wines with 11 variables on the chemical
properties of the wine. At least 3 wine experts rated the quality of each wine, 
providing a rating between 0 (very bad) and 10 (very excellent).

We will conduct an Exploratory Data Analysis in order to develop intuition
about this dataset, extract insights that may uncover relevant questions,
and eventually prepare the development of predictive models.

A first question that comes to mind is:

> Which chemical properties influence the quality of red wines?

We will start by conducting univariate analyses to identify variables that have 
little or no impact on wine quality, focusing on the variation of the variables.

Bivariate analyses will allow us to look deeper into the relationship between
retained variables and quality. This should enable us to identify critical
variables.

These critical variables will be further explored with multivariate analysis. We
should then be able to make predictions about wine quality based on its chemical
properties.



# Loading Packages and Dataset

```{r Packages, echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
library(gridExtra)
library(corrplot)
library(GGally)
library(scales)
library(memisc)
library(RColorBrewer)
```

```{r Dataset, echo=FALSE}
w = read.csv('wineQualityReds.csv')
```

# Univariate Analysis

## General information

First of all, let's get to know our dataset a little better.

Dimensions:
```{r Dimensions, echo=FALSE}
dim(w)
```

These are the names of our variables:

```{r Variables, echo=FALSE}
str(w)
```

We're going to rename the sulfur dioxide columns right away:
```{r Renaming Sulfur Variables}
names(w)[names(w) == 'free.sulfur.dioxide'] <- 'free.SO2'
names(w)[names(w) == 'total.sulfur.dioxide'] <- 'total.SO2'
```


Summary:
```{r Summary, echo=FALSE}
summary(w)
```

> Let's begin by looking at the distribution of our wines in terms of our
different variables

```{r Quality Structure, echo = FALSE}

str(w$quality)
str(w$fixed.acidity)
str(w$volatile.acidity)
str(w$citric.acid)
str(w$residual.sugar)
str(w$chlorides)
str(w$free.sulfur.dioxide)
str(w$total.sulfur.dioxide)
str(w$density)
str(w$ph)
str(w$sulphates)
str(w$alcohol)
```

## Dropping the X variable

The 'X' variable serves as an index, and we won't need it here. Let's get rid of
it right now so we don't have to subset our dataframe all along the analysis.

```{r Dropping the X variable}
w <- subset(w, select = - X)
```


## Variable distribution

To begin, we're going to look at the individual distribution of our variables.

```{r Variables Histograms - Logarithmic Correction | Quality, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(aes(x = quality), data = w) +
  geom_bar() +
  scale_x_continuous(breaks = seq(3, 8, by = 1)) +
  xlab('Quality') +
  ylab('Count') +
  ggtitle('Quality Distribution')

summary(w$quality)
```

In theory, grades can range from 0 to 10.
Effectively, they range from 3 to 8, with a median at 6 and a mean at 5.636.

Quality follows a normal distribution. As such, we have little data regarding
very low and very high grades, and mustb be cautious when drawing conclusions
from these.

```{r Variables Histograms - Logarithmic Correction | Fixed Acidity, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(aes(x = fixed.acidity), data = w) +
  geom_histogram() +
  scale_y_log10() +
  xlab('Fixed acidity') +
  ylab('Count (log10)') +
  ggtitle('Fixed Acidity Distribution')

summary(w$fixed.acidity)
```

Fixed acidity transformed logarithmically follows a normal distribution.
Most values range from 4.60 to about 14 g / dm^3, with a few between 14 and 16.
The mean is 8.32 and the median 7.9.

```{r Variables Histograms - Logarithmic Correction | Volatile Acidity, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = volatile.acidity), data = w) +
  geom_histogram() +
  scale_y_log10() +
  xlab('Volatile acidity') +
  ylab('Count (log10)') +
  ggtitle('Volatile Acidity Distribution')

summary(w$volatile.acidity)
```

Volatile acidity transformed logarithmically follows a normal distribution.
Values range from 0.12 to 1.58 g / dm^3, with a mean at 0.5278 and a median at
0.52.

```{r Variables Histograms - Logarithmic Correction | Citric Acid, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = citric.acid), data = w) +
  geom_histogram() +
  scale_y_log10() +
  xlab('Citric acid') +
  ylab('Count (log10)') +
  ggtitle('Citric Acid Distribution')

summary(w$citric.acid)
```

Although we can't call it right skewed because values remain relatively even
compared to one another, low citric wines are more numerous than high citric
wine. 
Values range from 0 to 1 g / dm^3, but values at 1 are outliers. The mean is 0.271 and
the median 0.26.
The range is short: this value might be negligeable.

```{r Variables Histograms - Logarithmic Correction | Residual Sugar, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = residual.sugar), data = w) +
  geom_histogram() +
  scale_y_log10() +
  xlab('Residual sugar') +
  ylab('Count (log10)') +
  ggtitle('Residual Sugar Distribution')

summary(w$residual.sugar)
```

For residual sugar we have a right skewed distribution with a few outliers above
10 g / dm^3.

Values range from 0.9 to 15.5 g / dm^3. The mean is 2.539 and the median 2.2.

```{r Variables Histograms - Logarithmic Correction | Chlorides, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = chlorides), data = w) +
  geom_histogram() +
  scale_y_log10() +
  xlab('Chlorides') +
  ylab('Count (log10)') +
  ggtitle('Chlorides Distribution')

summary(w$chlorides)
```

Chlorides follow a right-skewed distribution as well. They range from
0.012 to 0.611 g / dm^3, with three clusters. The range is extremely small,
the impact of this variable might be negligeable.


```{r Variables Histograms - Logarithmic Correction | Free Sulfur Dioxide, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = free.SO2), data = w) +
  geom_histogram() +
  scale_y_log10() +
  xlab('Free SO2') +
  ylab('Count (log10)') +
  ggtitle('Free Sulfur Dioxide Distribution')

summary(w$free.SO2)
```

Most wines have low free S02: the higher the Free S02 level, the less the count.
We have an outlier around 68 mh / dm^3.
Values range from 1 to 72 g / dm^3, with a mean at 15.87 and a median at 14.

```{r Variables Histograms - Logarithmic Correction | Total Sulfur Dioxide, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = total.SO2), data = w) +
  geom_histogram() +
  scale_y_log10() +
  xlab('Total SO2') +
  ylab('Count (log10)') +
  ggtitle('Total Sulfur Dioxide Distribution')

summary(w$total.SO2)
```

Once again, wines with a low total S02 level are more numerous, and the higher
the level, the less wine there is in our sample.
Values range from 6 to 289 mg / dm^3, with a mean at 46.47 and a median at 38.

```{r Variables Histograms - Logarithmic Correction | Density, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = density), data = w) +
  geom_histogram() +
  xlab('Density') +
  ylab('Count') +
  ggtitle('Density Distribution')

summary(w$density)
```

Density follows a normal distribution, ranging from 0.9901 to 1.0037 g / cm^3,
with a mean at 0.9967 and a median at 0.9968.
It is distributed over over a very small range, so this variable might be
negligeable too.

```{r Variables Histograms - Logarithmic Correction | pH, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = pH), data = w) +
  geom_histogram() +
  xlab('pH') +
  ylab('Count') +
  ggtitle('pH Distribution')

summary(w$pH)
```

pH follows a normal distribution ranging from 2.74 to 4.01, with a few outliers
around 2.75 and above 3.75. The mean is 3.311 and the median 3.31.

```{r Variables Histograms - Logarithmic Correction | Sulphates, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = sulphates), data = w) +
  geom_histogram() +
  scale_y_log10() +
  xlab('Sulphates') +
  ylab('Count (log10)') +
  ggtitle('Sulphates Distribution')

summary(w$sulphates)
```

Sulphates follows a normal distribution as well, a little skewed to the right.
We have outliers around 1.6 and 1.8 g / dm^3. The values range from 0.33 to
2 g / dm^3, with a mean at 0.6581 and a median at 62.

```{r Variables Histograms - Logarithmic Correction | Alcohol, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(aes(x = alcohol), data = w) +
  geom_histogram() +
  xlab('Alcohol') +
  ylab('Count') +
  ggtitle('Alcohol Distribution')

summary(w$alcohol)
```

The alcohol distribution is right skewed, ranging from 8.40 degrees by volume to
14.90. We have outliers below 9, and above 14. The mean is 10.42 and the median
10.20.

## Structure of the dataset

- Although the quality column should hold values between 0 and 10, in reality
our values range from 3 to 8
- Alcohol values range from 8.40 to 14.90
- Free and total sulfur dioxide, fixed and volatile acidity, as well as citric
acid, are distributed over a wide range of values
- Density and pH seem to follow a normal distribution
- Fixed acidity, volatile acidity, residual sugar and chlorides seem to have
extreme outliers.
- Free sulfur dioxide, total sulfur dioxide and sulphates are right skewed.


## Insights and features of interest

The main feature of interest of this dataset is the quality variable, which is
supposedly impacted by all the other variables.

Most of our variables were highly right skewed, so we had to use logarithmic
transformations.

Density and chlorides are distributed over very small ranges. No matter
the expertise of the three oenologists that graded the wines, it would be
unimaginable to distinguish variations over such a small range. Therefore, it is
likely that these variables had a negligeable impact over the final quality
values.

Fixed acidity and alcohol, on the other hand, may very well have an important
weight in the final grade.

Finally, because the quality histogram follows a gaussian distribution, we
should be cautious regarding our our analyses and conclusions about low and high
quality wines.


# Bivariate analysis

First of all, we're going to generate a correlation matrix, to gain general
insights about the relationship between all of our variables.

## Exploring correlations

```{r Correlation Matrice, echo=FALSE}
res <- cor(w)
corrplot(res, type = 'lower', tl.srt=45, title = 'Correlation Matrice', mar=c(0,0,1,0))
```

Positive correlations are blue, negative correlations are red.
Strong correlations are big and dark, weak correlations are small and light.

Here, we see that quality seems to be strongly tied to alcohol and volatile
acidity, and to a lesser extent to sulphates and citric acid.

For the sake of our general wine chemical properties erudition, let's also note
that:
- density is strongly correlated to fixed acidity and alcohol
- pH is strongly correlated to fixed acidity and citric acid

It makes sense that citric acid, volatile acidity and fixed acidity are
correlated, as well as free and total sulfur dioxide:

```{r Scatter plot of Free SO2 vs Total SO2, echo=FALSE}
ggplot(aes(x = free.SO2, 
           y = total.SO2), 
       data = w) +
      geom_point(alpha = 0.3, size = 1) +
      geom_smooth(method = "lm", se = FALSE,size=1)  +
  scale_color_brewer(type='seq',
                   guide=guide_legend(title='Quality')) +
  ggtitle('Scatter plot of Free SO2 vs Total SO2')
```

```{r Correlation between Free SO2 and Total SO2, echo=FALSE}
cor.test(w$free.SO2, w$total.SO2)
```

Total Sulfur Dioxide is made up of the amount of free and bound forms of SO2,
so it makes sense to witness a strong regular correlation between both.

```{r Scatter plot of Citric Acid vs Volatile Acidity, echo=FALSE}
ggplot(aes(x = citric.acid, 
           y = volatile.acidity), 
       data = w) +
      geom_point(alpha = 0.3, size = 1) +
      geom_smooth(method = "lm", se = FALSE,size=1)  +
  scale_color_brewer(type='seq',
                   guide=guide_legend(title='Quality')) +
  ggtitle('Scatter plot of Citric Acid vs Volatile Acidity')
```

```{r Correlation between Citric Acid vs Volatile Acidity, echo=FALSE}
cor.test(w$citric.acid, w$volatile.acidity)
```

Citric acid is a fixed acid, and adds freshness and flavor to the wine. On the
other hand, volatile acidity corresponds to the amount of acetic acid, which can
lead to unpleasant vinegar taste in high levels. It makes sense the more of a
fixed acid we find in a wine, the less volatile acid there is.

```{r Scatter plot of Citric Acid vs Fixed Acidity, echo=FALSE}
ggplot(aes(x = citric.acid, 
           y = fixed.acidity), 
       data = w) +
      geom_point(alpha = 0.3, size = 1) +
      geom_smooth(method = "lm", se = FALSE,size=1)  +
  scale_color_brewer(type='seq',
                   guide=guide_legend(title='Quality')) +
  ggtitle('Scatter plot of Citric Acid vs Fixed Acidity')
```

Echoing what we just wrote and witnessed, it also makes sense that the more
citric acid there is, the higher the level of fixed acidity. Of course, the 
correlation is trong but not perfect, since there are other fixed acids in wine
(malic acid, tartaric acid...)

```{r Correlation between Citric Acid vs Fixed Acidity, echo=FALSE}
cor.test(w$citric.acid, w$fixed.acidity)
```

```{r Scatter plot of Volatile Acidity vs Fixed Acidity, echo=FALSE}
ggplot(aes(x = volatile.acidity, 
           y = fixed.acidity), 
       data = w) +
      geom_point(alpha = 0.3, size = 1) +
      geom_smooth(method = "lm", se = FALSE,size=1)  +
  scale_color_brewer(type='seq',
                   guide=guide_legend(title='Quality')) +
  ggtitle('Scatter plot of Volatile Acidity vs Fixed Acidity')
```

```{r Correlation between Volatile Acidity vs Fixed Acidity, echo=FALSE}
cor.test(w$volatile.acidity, w$fixed.acidity)
```

And to finish stating the obvious, the higher the level of volatile acidity,
the lower the level of fixed acidity.

## Validating insights with further bivariate analysis

```{r Boxplot of Quality vs Alcohol, echo=FALSE}

ggplot(aes(factor(quality), 
            alcohol), 
        data = w) +
  geom_jitter( alpha = .3)  +
  geom_boxplot( alpha = .5,color = 'blue')+
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4) +
  ggtitle('Boxplot of Quality vs Alcohol')
```

```{r Correlation between Quality and Alcohol, echo=FALSE}
cor.test(w$quality, w$alcohol)
```

```{r Statistics for Quality vs Alcohol, echo=FALSE}
by(w$alcohol, w$quality, summary)
```

There's a moderately strong correlation between quality and alcohol: 0.476

Here, we can see that our quality grade definitely goes up with the alcohol rate
medians. It would indicate that alcohol has an important impact on quality.

We can see here that quality seems increase with the rate of alcohol:
- For the grade 5, the second and third quartiles are between 9.40 and 10.20 
degrees of alcohol, and the median is 9.70
- For the grade 6, they are between 9.80 and 11.30, and the median is 10.50
- For the grade 7, they are between 10.80 and 12.10, and the median is 11.50
- For the grade 8, they are between 11.32 and 12.88, and the median is 12.15

```{r Boxplot of Quality vs Volatile Acidity, echo=FALSE}

ggplot(aes(factor(quality), 
            volatile.acidity), 
        data = w) +
  geom_jitter( alpha = .3)  +
  geom_boxplot( alpha = .5,color = 'blue')+
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4) +
  ggtitle('Boxplot of Quality vs Volatile Acidity')
```

```{r Correlation between Quality and Volatile Acidity, echo=FALSE}
cor.test(w$quality, w$volatile.acidity)
```

```{r Statistics for Quality vs Volatile Acidity, echo=FALSE}
by(w$volatile.acidity, w$quality, summary)
```

There's a moderately strong inverse correlation between quality and volatile
acidity: -0.391

We can see here that quality seems increase with the rate of alcohol:
- For the grade 3, the second and third quartiles are between 0.6475 and 1.01 
degrees of alcohol, and the median is 0.845
- For the grade 4, they are between 0.53 and 0.87, and the median is 0.67
- For the grade 5, they are between 0.46 and 0.67, and the median is 0.58
- For the grade 6, they are between 0.38 and 0.60, and the median is 0.49
- For the grade 7, they are between 0.30 and 0.485, and the median is 0.37
- For the grade 8, they are between 0.335 and 0.4725, and the median is 0.37

```{r Boxplot of Quality vs Sulphates, echo=FALSE}

ggplot(aes(factor(quality), 
            sulphates), 
        data = w) +
  geom_jitter( alpha = .3)  +
  geom_boxplot( alpha = .5,color = 'blue')+
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4) +
  ggtitle('Boxplot of Quality vs Sulphates')
```

```{r Correlation between Quality and Sulphates, echo=FALSE}
cor.test(w$quality, w$sulphates)
```

```{r Statistics for Quality vs Sulphates, echo=FALSE}
by(w$sulphates, w$quality, summary)
```

There's a small correlation between quality and sulphates level: 0.251

Once again we can see a correlation, although it is definitely weaker than
what we witnessed before:
- For the grade 3, the second and third quartiles are between 0.5125 and 0.6150 
degrees of alcohol, and the median is 0.5450
- For the grade 8, the second and third quartiles are between 0.6900 and 0.8200 
degrees of alcohol, and the median is 0.7400

## Insights and features of interest

We've found that quality is correlated with alcohol and sulphates, and inversely
correlated with volatile acidity.

We've confirmed that in general, the higher the rate of alcohol, the
higher the rate of sulfates and the lower the level of volatile acidity, the
better the grade.

The relationship is especially strong between quality and alcohol, and extremely
regular with volatile acidity.

We've also seen that:
- free sulfur dioxide is positively correlated with total sulfur dioxide
- citric acid is positively correlated with fixed acidity
- citric acid is inversely correlated with volatile acidity
- fixed acidity is inversely correlated with volatile acidity

# Multivariate analysis

## Scatter plot of Quality by Alcohol and Volatile Acidity

Let's first focus our multivariate analysis about quality on the alcohol and
volatile acidity variables, since our bivariate analysis showed they seem to be
the more impactful.

We will omit the top 1% of the volatile acidity values to eliminate outliers, as
depicted below:

```{r Detecting Volatile Acidity Outliers}
ggplot(w, aes( x = 1, y = volatile.acidity)) + 
  geom_jitter(alpha = 0.1 ) +
  geom_boxplot(alpha = 0.2, color = 'red' )
```

Most of our data has a volatile acidity between 0.4 and a little above 0.6.

```{r Quality vs Volatile Acidity and Alcohol, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(aes(x = volatile.acidity, y = alcohol, color = as.factor(quality)),
       data = w) + 
  geom_point(alpha = 1, size = 1, position = 'jitter') +
  geom_smooth(method = 'lm', se = FALSE, size = 1) +
  scale_x_continuous(lim = c(0, quantile(w$volatile.acidity, 0.99))) +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Quality', reverse = T,
                                          override.aes = list(alpha = 1,
                                                              size = 2))) +
  ggtitle('Quality by Volatile Acidity and Alcohol') +
  xlab('Volatile acidity level') +
  ylab('Alcohol (%)')
```

We can identify a cluster here, loosely in the 11 to 13 range for alcohol
degree, and 0.2 and 0.4 for volatile acidity, where dots tend to be high quality
green. The higher the volatile acidity, the hotter the color. The same holds
true for low alcohol levels, where yellow dominates.


## Scatter plot of Quality by Alcohol and Sulphates

Sulphates seemed to be another impactful variable. Let's plot it against alcohol
and quality.

```{r Scatter Plot of Quality vs Volatile Acidity and Alcohol, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(aes(x = sulphates, y = alcohol, color = as.factor(quality)), data = w) + 
  geom_point(alpha = 1, size = 1, position = 'jitter') +
  geom_smooth(method = 'lm', se = FALSE, size = 1) +
  scale_x_continuous(limits = c(0, 1.5), breaks = seq(0, 1.5, 0.1)) +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Quality', reverse = T,
                                          override.aes = list(alpha = 1,
                                                              size = 2))) +
  ggtitle('Quality by Alcohol and Sulphates') +
  xlab('Sulphate level') +
  ylab('Alcohol (%)')
```

Once again, we can identify a cluster where the highest quality wines have an
alcohol rate loosely in the 11 to 13 range, and a sulphate level between 0.6
and 0.9.

## Insights

Quality seems to be impacted at the same time by alcohol rate, volatile acidity
and sulphates levels.


## Laying down the foundations for a model

```{r Building the Loess Model}
summary(loess(I(quality) ~ I(volatile.acidity+ alcohol), data = w))
```

# Final Plots and Summary

## Wrap up and final plots

This dataset contains 1,599 red wines with 11 variables on the chemical
properties of the wine.

Univariate analysis enabled us to understand the distribution of each variable,
and to eliminate chlorides and density as impactful variables.

```{r First Final Plot, echo=FALSE, message=FALSE, warning=FALSE}

p1 <- ggplot(aes(x = chlorides), data = w) +
  geom_histogram(fill = I('#0080ff')) +
  xlab('Chlorides') +
  ylab('Count') +
  ggtitle('Distribution of Chlorides')

p2 <- ggplot(aes(x = density), data = w) +
  geom_histogram(fill = I('#0080ff')) +
  xlab('Density') +
  ylab('Count') +
  ggtitle('Distribution of Density')

grid.arrange(p1, p2, ncol = 2)
```

Our bivariate analysis allowed us to identify chlorides and density as
negligeable variables, because of their short range implying a difficulty to
distinguish a real impact. Alcohol, volatile acidity and
sulphates, on the other hand, were identified as potentially being the most
impactful variables on quality. The first two follow a normal distribution like
the quality variable.

```{r Second Final Plot, echo=FALSE, message=FALSE, warning=FALSE}

ggplot(aes(factor(quality), 
            volatile.acidity), 
        data = w) +
  geom_jitter( alpha = .3)  +
  geom_boxplot( alpha = .5,color = 'blue')+
  stat_summary(fun.y = "mean", 
               geom = "point", 
               color = "red", 
               shape = 8, 
               size = 4) +
  ggtitle('Boxplot of Quality vs Volatile Acidity')
```

Our bivariate analysis confirmed the insights brought out by our univariate
analysis. When quality goes up, the volatile acidity median goes down, and the
alchohol median goes up. The sulphates median goes up as well, although to a
lesser extent.

Finally, our multivariate analysis confirmed our conjecture.
We hypothesized that quality was linked to a high enough degree of alcohol,
a low degree of volatile acidity and possibly, a high enough level of sulphates.

```{r Third Final Plot, echo=FALSE, message=FALSE, warning=FALSE}

p1 <- ggplot(aes(x = volatile.acidity, y = alcohol, color = as.factor(quality)),
             data = w) + 
  geom_point(alpha = 1, size = 1, position = 'jitter') +
  geom_smooth(method = 'lm', se = FALSE, size = 1) +
  scale_x_continuous(lim = c(0, quantile(w$volatile.acidity, 0.99))) +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Quality', reverse = T,
                                          override.aes = list(alpha = 1,
                                                              size = 2))) +
  ggtitle('Quality by Volatile Acidity and Alcohol') +
  xlab('Volatile acidity level') +
  ylab('Alcohol (%)')

p2 <- ggplot(aes(x = sulphates, y = alcohol, color = as.factor(quality)),
             data = w) + 
  geom_point(alpha = 1, size = 1, position = 'jitter') +
  geom_smooth(method = 'lm', se = FALSE, size = 1) +
  scale_x_continuous(lim = c(0, quantile(w$sulphates, 0.99))) +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Quality', reverse = T,
                                          override.aes = list(alpha = 1,
                                                              size = 2))) +
  ggtitle('Quality by Alcohol and Sulphates') +
  xlab('Sulphate level') +
  ylab('Alcohol (%)')

grid.arrange(p1, p2, ncol = 1)
```

Our analysis clearly showed that there is a cluster of good quality wines,
loosely in the 11 to 13 range for alcohol degree, 0.2 and 0.4 g / dm^3 for
volatile acidity, and a sulphate level between 0.6 and 0.9 g / dm3.

# Final reflections

## Insights

Alcohol rate, volatile acidity and sulphate levels all impact the final grade a 
wine receives from a pannel of experts.

## Limitations

We have to keep in mind the limitations of this model.

The dataset only contains 1,599 red wines. It is in no way representative of all
the red wines across the world. Other variables may impact the quality of the
wine: its preservation, its origin, its age, its cépage...

We also stated that we have very few values regarding low and high quality wine,
so our conclusions must be taken cautiously.

## Future work

I'd be interested in working with a dataset with additional properties, as
written above: age, cépage, origin...

It would also be great to conduct a similar analysis about white wine, to see if
the impactful variables are the same. Then, an analysis comparing red and white
wine would be interesting.

And being from Reims, I'm definitely going to look for a Champagne dataset!

## Struggles

At first, I struggled about how to build the analysis from A to Z. I also found
it difficult to identify the impactful variables from the univariate section.
Plotting the multivariate section wasn't easy either, and although I hinted that
we could start building a model, I actually don't have a clue as to how to do it
using another method than linear regression.