---
title: "Stroop Effect Analysis"
output:
  html_document: default
---


## Introduction
In a Stroop task, participants are presented with a list of words, with each 
word displayed in a color of ink. The participant’s task is to say out loud the
color of the ink in which the word is printed. The task has two conditions: a
congruent words condition, and an incongruent words condition. In the congruent
words condition, the words being displayed are color words whose names match the
colors in which they are printed: for example <span style="color: #C70039">RED</span>, <span style="color: #39A6E4">BLUE</span>. In the incongruent words condition, the words displayed are color words whose names do not match the colors in which they are printed: for example <span style="color: #E8A12B">PURPLE</span>, <span style="color: #82E439">ORANGE</span>. In each case, 
we measure the time it takes to name the ink colors in equally-sized lists. Each
participant will go through and record a time from each condition.

Our dataset is a sample of a larger population, made up of 24 individuals.


## Dependent and independent variables
Our *independent* variable is the **color matching** (does the color match the
word or not).
Our *dependent* variable is the **time it takes to name the ink color**.


## Null and alternative hypotheses

### Null and alternative hypothesis in words
We can hypothesize that **the distribution will be normal**: the time to
name should fall around a certain value in both conditions.
However, we can also hypothesize that **it will take longer to name the color** 
**in an incongruent condition**, because the situation is more confusing.

* Our *null hypothesis* is that **there is no difference in population means of 
response time under incongruent and congruent scenarios**.
* Our *alternative hypothesis* is that **the population mean of response time
under the incongruent scenari is higher than the population mean of response 
time under the congruent scenari**.

We are going to test these hypotheses to check if there is enough evidence 
to infer that the condition is true for the entire population.

### Null and alternative hypothesis in mathematical terms
In mathematical terms, let's define:

* $\mu C$ as the population mean of the **Congruent** sample
* $\mu I$ as the population mean of the **Incongruent** sample

$\ H_0: \mu C = \mu I$
$\ H_1: \mu C < \mu I$

## Descriptive statistics

#### Summary statistics
```{r Summary statistics}
library(ggplot2)
library(reshape2)

stroop = read.csv('stroopdata.csv')
summary(stroop)
```


#### IQR for Congruent colors
```{r IQR for Congruent colors}
IQR(stroop$Congruent)
```


#### IQR for Incongruent colors
```{r IQR for Incongruent colors}
IQR(stroop$Incongruent)
```


#### Variance for Congruent colors
```{r Variance for Congruent colors}
var(stroop$Congruent)
```


#### Variance for Incongruent colors
```{r Variance for Incongruent colors}
var(stroop$Incongruent)
```


#### Standard deviation for Congruent colors
```{r Standard deviation for Congruent colors}
sd(stroop$Congruent)
```


#### Standard deviation for Incongruent colors
```{r Standard deviation for Incongruent colors}
sd(stroop$Incongruent)
```


## Visualization

### Histograms and density kernels

```{r Histogram and density kernel for Congruent colors}

ggplot(stroop, aes(x=Congruent)) + 
    geom_histogram(aes(y=..density..),
                   binwidth=2,
                   colour="black", fill="#ADD8E6") +
    geom_density(alpha=.5, fill="#FFE4B5") +
    xlab('seconds') +
    ggtitle('Distribution for congruent colors')
```

The distribution for congruent colors is **normal** as expected.

```{r Histogram and density kernel for Incongruent colors}
ggplot(stroop, aes(x=Incongruent)) + 
    geom_histogram(aes(y=..density..),
                   binwidth=2,
                   colour="black", fill="#ADD8E6") +
    geom_density(alpha=.5, fill="#FFE4B5") +
    xlab('seconds') +
    ggtitle('Distribution for incongruent colors')
```

The distribution for incongruent colors is actually right skewed. We could
consider these values as outsiders, but we should consider the fact that
incongruent colors really confuses a small part of the population.

### Boxplots

```{r Boxplots}
boxplot(stroop, staplelty=0, col="#ADD8E6", main="Boxplots of response time per condition", 
  	xlab="Color matching", ylab="Seconds to answer")
```

From the boxplots alone, it does **seem like incongruent color matching results
in longer response time overall**. The two higher values we discussed above while
commenting the Incongruent distribution are considered as **outsiders** here.

## Expected statistical test

We must choose our statistical test according to our situation.
We want to know if, in general, it takes longer to name a color when the word is
incongruent, compared to when the word is congruent. Therefore, we need to 
compare means or medians.
Here are the other information at our disposal:
  
* we have **two samples**
* **means and medians for each condition are close**, so the mean is a good measure
of comparison
* we **don't know** the **population standard deviation**
* our samples are **normally distributed** (we consider the high values for the 
incongruent scenari as outliers).

Because we don't know the popoluation standard deviation, and because our 
samples contain less than 30 individuals, a t-test is preferable to a
z-test. Also, because the samples are normally distributed, a t-test is 
preferable to a Wilcoxon signed-rank test.

The **t-test** will allow us to determine if the two sets of data are
significantly different from each other. We will actually
use a **paired t-test**: this is perfect to compare two samples before and after
a specific event (here, switching from congruent to incongruent colors).
We will make it **one-tailed**, because we suspect the  incongruent variable
makes reading slower, not faster.

## Paired t-test

```{r t-test}
t.test(stroop$Congruent, stroop$Incongruent, paired=TRUE)
```

Our **p-value is less than 0.05**, we **reject our null hypothesis** (the idea that 
there is no relation between color congruence and time). We can conclude 
that **there is a definite correlation between reading time and color congruence.**


## Conclusion

I believe the confusion between the meaning of the word and its actual color is
responsible for the effect observed. If the words written are not colors but 
animals or fruits, then the reading process is smoother. A somehow similar
effect may be obtained by **matching a direction to a movement instead of an
orientation**. If the arrow is pointed up but moves to the right, the subject
should swipe right, but it's not so intuitive.