---
title: "Diamonds Price EDA and Prediction"
output:
  html_document: default
  html_notebook: default
---

The 'diamonds' dataset is one of the datasets provided with the ggplot2 R
package. We're going to see if we can predict the price of a diamond based on
its characteristics.

We will first conduct an EDA to get to know the data and analyze the impact of
the different variables. We will then push the analysis further in order to
build a linear model and use it to predict prices.

NOTE: This notebook contains two analyses condensed in one. While they
definitely are related, there are a few points that need to be revised to help
the reader follow the logic easily. This will be done in the near future, but
we found that this notebook was interesting enough to be published in its raw
form, waiting for revision.

# Loading packages and dataset

```{r Loading packages and dataset, warning=FALSE}
library(ggplot2)
library(GGally)
library(scales)
library(memisc)
library(RColorBrewer)

data("diamonds")
```

# Univariate Analysis

## General Information

Dimensions of the dataset:
```{r Dimensions}
dim(diamonds)
```

Name of the variables:
```{r Name of the variables}
str(diamonds)
```

Summary:
```{r Summary}
summary(diamonds)
```

Levels of our cateogricalvariables:
```{r Levels}
levels(diamonds$cut)
levels(diamonds$color)
levels(diamonds$clarity)
```

# Univariate analysis

Let's jump right into it and focus our analysis on price.

## Price histogram
```{r Price histogram}
qplot(data = diamonds, x = price)
summary(diamonds$price)
```

The distribution of diamonds prices is clearly right-skewed.


## Diamond price detail

Let's get some numbers:

```{r # Diamond price detail}
sum(diamonds$price < 500)
sum(diamonds$price < 250)
sum(diamonds$price >= 15000)
```

Our dataset contains:
- 1729 diamonds with a price below $500
- 0 diamonds with a price below $250
- 15,000 diamonds with a price equal to or above $15,000

## Histogram - cheaper diamonds

Let's get a look at the cheapest diamonds:

```{r}
qplot(data = diamonds, x = price,
      binwidth = 20) +
  scale_x_continuous(limits = c(0, 1500), breaks = seq(0, 1500, 100))

Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

Mode(diamonds$price)
```

The mode of the cheapest diamonds (with a price between $0 and $1,500) is 605.

# Bivariate analysis

## Faceting - Histogram of diamond prices by cut

Let's facet our prices by the quality of the cut.

## Scaling - Histogram of diamond prices by cut
```{r}
qplot(data = diamonds, x = price) +
  facet_wrap(~cut, ncol= 2, scales = 'free_y')

by(diamonds$price, diamonds$cut, summary)
```

## Faceting - Histogram of price per carat by cut
```{r}
qplot(data = diamonds, x = price / carat, binwidth = 0.1) +
  facet_wrap(~cut) +
  scale_x_log10()
```

The price per carat definitely seems to increase with the quality of the cut.

## Price boxplots and statistics
```{r}
qplot(data = diamonds, x = clarity, y = price, geom = 'boxplot')
by(diamonds$price, diamonds$clarity, summary)
```

## Color boxplots and statistics
```{r}
qplot(data = diamonds, x = color, y = price, geom = 'boxplot')
by(diamonds$price, diamonds$color, summary)
```

## Faceting - Boxplots of price per color by cut
```{r}
qplot(data = diamonds, x = color, y = price / carat, geom = 'boxplot') +
  coord_cartesian(ylim = c(250, 6000))
```

## Frequency polygon - Carat
```{r}
qplot(data = diamonds, x = carat, binwidth = 0.01, geom = 'freqpoly') +
  scale_x_continuous(limits = c(0, 1.5), breaks = seq(0, 1.5, 0.1))

```

# Scatter plot - Price vs x
```{r}
ggplot(aes(x = x , y = price), data = diamonds) +
  geom_jitter(alpha= 1/20) +
  xlim(3, 9)
```

## Correlation between Price and x, y and z
```{r}
cor.test(diamonds$price, diamonds$x)
cor.test(diamonds$price, diamonds$y)
cor.test(diamonds$price, diamonds$z)
```

## Scatter plot of price vs depth
```{r}
ggplot(aes(x = depth, y = price), data = diamonds) +
  geom_point(alpha = 0.05,
             position = position_jitter(h = 0),
             color = 'orange')
```

## Correlation between Price and depth
```{r}
cor.test(diamonds$price, diamonds$depth)
```

## Scatter plot of price vs carat omitting the top 1% percentile
```{r}
ggplot(aes(x = carat, y = price), data = diamonds) +
  geom_point() + 
  xlim(0, quantile(diamonds$carat, 0.99)) +
  ylim(0, quantile(diamonds$price, 0.99))
```

## Scatter plot of price vs volume (x * y * z)
```{r}
diamonds$volume <- diamonds$x * diamonds$y * diamonds$z
ggplot(aes(x = volume, y = price), data = diamonds) +
  geom_point()
```

## Correlation between price and volume
```{r}

with(subset(diamonds, volume > 0 & volume < 800), cor(price, volume))
```

## Adjustment - price vs volume
```{r}
ggplot(aes(x = volume, y = price), data = subset(diamonds, volume > 0 & volume < 800)) +
  geom_point(alpha = 1/20) +
  geom_smooth(method = 'lm', color = 'red')
```

## Mean Price by Clarity
```{r}
library(dplyr)
diamondsByClarity <- diamonds %>%
  group_by(clarity) %>%
  summarize(mean_price = mean(price),
            median_price = median(price),
            min_price = min(price),
            max_price = max(price),
            n = n()) %>%
  arrange(clarity)
head(diamondsByClarity)
```

## Bar Charts of Mean Price
```{r}
diamonds_by_clarity <- group_by(diamonds, clarity)
diamonds_mp_by_clarity <- summarise(diamonds_by_clarity, mean_price = mean(price))

diamonds_by_color <- group_by(diamonds, color)
diamonds_mp_by_color <- summarise(diamonds_by_color, mean_price = mean(price))

p1 <- ggplot(aes(x = clarity, y = mean_price), data = diamonds_mp_by_clarity) +
  geom_bar(stat = 'identity')

p2 <- ggplot(aes(x = color, y = mean_price), data = diamonds_mp_by_color) +
  geom_bar(stat = 'identity')

library(gridExtra)
grid.arrange(p1, p2, ncol = 1)
```

There's something very odd here. It goes against the intuition that price goes
down when color and clarity are better. Maybe there's another variable
influencing this behavior. Let's look at the influence of the cut.

```{r}
diamonds_by_cut <- group_by(diamonds, cut)
diamonds_mp_by_cut <- summarise(diamonds_by_cut, mean_price = mean(price))

p3 <- ggplot(aes(x = cut, y = mean_price), data = diamonds_mp_by_cut) +
  geom_bar(stat = 'identity')

grid.arrange(p1, p2, p3, ncol = 1)
```


#Predictions

## Scatterplot Review

```{r Scatterplot Review}
library(ggplot2)
data(diamonds)

ggplot(aes(x = carat, y = price), data = diamonds) +
  geom_point(fill = I('#F79420'), color = I('black'), shape = 21) +
  xlim(0, quantile(diamonds$carat, .99)) +
  ylim(0, quantile(diamonds$price, .99))
```


## Price and Carat Relationship

- We can see a non-linear relationship. Maybe it's exponential, maybe it's
something else.
- We can see that the dispersion or variance also increases while carat size
increases.

## Add on a Linear Model

```{r Add on a Linear Model, message=FALSE, warning=FALSE}
ggplot(aes(x = carat, y = price), data = diamonds) +
  geom_point(color = '#F79420', alpha = 1/4) +
  stat_smooth(method = 'lm') +
  scale_x_continuous(lim = c(0, quantile(diamonds$carat, 0.99))) +
  scale_y_continuous(lim = c(0, quantile(diamonds$price, 0.99)))
```

## ggpairs Function

```{r ggpairs Function, message=FALSE, warning=FALSE}

# sample 10,000 diamonds from the data set
set.seed(20022012)
diamond_samp <- diamonds[sample(1:length(diamonds$price), 10000), ]
ggpairs(diamond_samp,
        axisLabels = 'internal',
        lower = list(continuous = wrap('points', shape = I('.'))),
        upper = list(combo = wrap('box', outlier.shape = I('.'))))
```

Here is what we can see from the matrice.
In the lower triangle, Ggplot uses:
- grouped histograms for qualitative / qualitative pairs
- scatter plots for quantitative / quantitative pairs

In the upper triangle, Ggplot uses:
- grouped histograms for qualitative / qualitative pairs, this time using the x
instead of the y variable as the grouping factor
- box plots for qualitative / quantitative pairs
- correlation coefficients for quantitative / quantitative pairs

We can see what might be relationship between price and clarity and price and
color, which we will keep in mind for later when we'll start modeling our data.

The critical factor driving price is the size, or the carat weight of the
diamond. As we saw before, the relationship between price and diamond size is
nonlinear. What might explain this pattern? On the supply side, larger
continuous chunks of diamonds without significant flaws are probably harder to
find than smaller ones. This might explain this sort of exponential looking
curve.

This is related to the fact that the weight of a diamond is a function of
volume, and volume is a function of the length times the width times the height
of a diamond. This suggests that we might be especially interested in the cube
root of carat wieght.

It's often the case that leveraging substantive knowledge about your data like
this can lead to especially fruitful transformations.


## The Demand of Diamonds

On the demand side, customers in the market for a less expensive, smaller
diamond are probably more sensitive to price than more well-to-do buyers.
Many less than one carat customers would surely never buy a diamond, were it
not for the social norm of presenting one when proposing.

There are fewer customers who can afford a bigger diamond that is larger than
one carat, hence we shouldn't expect the market for bigger diamonds to be as
competitive as the one for smaller diamonds. So it makes sense that the variance
as well as the price would increase with carat size.

Often, the distribution of any monetary variable like dollars will be highly
skewed and vary over orders of magnitude. This can result from path dependence
(the rich getting richer), or multiplicative processes like year on year
inflation, or some combination of both.

Hence it's a good idea to look into compressing any such variable by putting it
on a log scale.

```{r The Demand of Diamonds}
library(gridExtra)

plot1 <- qplot(data = diamonds, x = price, binwidth = 100, fill = I('#099DD9')) + 
  ggtitle('Price')

plot2 <- qplot(data = diamonds, x = price, binwidth = .01, fill = I('#F79420')) +
  ggtitle('Price (log10)') +
  scale_x_log10()

grid.arrange(plot1, plot2)
```


## Connecting Demand and Price Distributions

We can see that the prices for diamonds are pretty heavily skewed.

But when we put thos prices on a log10 scale, they seem much better behaved:
they are much closer to the bell curve of a normal distribution. We can even see
a little bit of evidence of bimodality on this log10 scale, which is consistent
with our two class rich buyer and poor buyer speculation about the nature of
customers for diamonds. 


## Scatterplot Transformation

```{r Scatterplot Transformation}
qplot(carat, price, data = diamonds) +
  scale_y_continuous(trans = log10_trans()) +
  ggtitle('Price (log10) by Carat')
```

This plot looks better than before. On the log 10 scale, the prices look less
dispersed at the high end of Carat size and price, but we can do better.

We're trying to use the cube root of Carat in light of our speculation about
flaws being exponentially more likely in diamonds with more volume (volume is on
a cubic scale).

First, we need a function to transform the Carat variable.

## Create a new function to transform the carat variable

```{r cuberoot transformation}
cuberoot_trans = function() trans_new('cuberoot', transform = function(x) x^(1/3),
                                      inverse = function(x) x^3)
```

## Use the cuberoot_trans function
```{r Use cuberoot_trans, message=FALSE, warning=FALSE}
ggplot(aes(carat, price), data = diamonds) + 
  geom_point() + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')
```

We can see that with these transformations we used to get our data on this
scale, things look almost linear; We can now move forward and see about
modelling our data using just a linear model.


## Overplotting Revisited

So far, we haven't done anything about overplotting (when multiple points take
on the same value).

```{r Sort and Head Tables}
head(sort(table(diamonds$carat), decreasing = T))
head(sort(table(diamonds$price), decreasing = T))
```

As we can see, we have a vast amount of points at the same price, which will
result in some serious overplotting. This can obscure the density and the
sparsity of our data at really key points. We can deal with this by making our
points smaller, jittering them, and adding transparency.

```{r Overplotting Revisited, message=FALSE, warning=FALSE}
ggplot(aes(carat, price), data = diamonds) + 
  geom_point(alpha = 0.5, size = 0.75, position = 'jitter') +
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')
```


### Other Qualitative Factors

We can see what looks like an almost linear relationship between carat weight
and price after doing some transformations. But surely there are other factors
that must influence the price of a diamond.

Clarity seem to factor into price. However, many consumers are looking for a
diamond of a minimum size, so we shouldn't expect clarity to be as strong a
factor as carat weight.

According to Blue Nile, the cut of a diamond has a much more consequential
impact on that fiery quality that jewelers describe when they talk about
diamonds.

Many of the imperfections on clarity are microscopic and do not affect the
diamonds beauty in any discernible way.

Let's see if clarity, cut or color can explain some of the variants in price
when we visualize it on our plot using color.


## Price vs. Carat and Clarity

```{r Price vs. Carat and Clarity, message=FALSE, warning=FALSE}

ggplot(aes(x = carat, y = price, color = clarity), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
    guide = guide_legend(title = 'Clarity', reverse = T,
    override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
    breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
    breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Clarity')
```


## Clarity and Price

Clarity does seem to explain an awful lot of the remaining variance in price,
after adding color to our plot. Holding carat weight constant, looking at one
part of the plot, we the diamonds with lower clarity are almost always cheaper
than diamonds with better clarity.


## Price vs. Carat and Cut

Alter the code below.
```{r Price vs. Carat and Cut, message=FALSE, warning=FALSE}
ggplot(aes(x = carat, y = price, color = cut), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Cut', reverse = T,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Cut')
```


## Cut and Price

Despite what Blue Nile says, we don't see much variation on the cut. Most of the
diamonds in the data are ideal cut anyway, so we've lost the color pattern that
we saw before.


## Price vs. Carat and Color

Alter the code below.
```{r Price vs. Carat and Color, message=FALSE, warning=FALSE}
ggplot(aes(x = carat, y = price, color = color), data = diamonds) + 
  geom_point(alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Color', reverse = FALSE,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Color')
```


## Color and Price

Color does seem to explain some of the variance in price, just like the clarity
variable. Blue Nile however states that the difference between all color grades
from D to J are basically not noticeable to the naked eye. Yet, we do see the
color difference in the price tag.


## Linear Models in R

In R, we can create models using the lm function. lm(y~x) where y is the outcome
variable and x the explanatory variable.

Here, we would use:
log10(price) ~ carat^(1/3)
Price is the outcome and carat the predictor variable. Using our domain specific
knowledge of diamonds and carat weight, we know we must take the cube root of
carat weight (volume).

We apply a log transformation to our long tailed dollar variable, and we
speculate that the flawless diamond should become exponentially rarer as the
volume increases.


## Building the Linear Model

```{r Building the Linear Model}
m1 <- lm(I(log(price)) ~ I(carat^(1/3)), data = diamonds)
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)
mtable(m1, m2, m3, m4, m5, sdigits = 3)
```

Notice how adding cut to our model does not help explain much of the variance
in the price of diamonds. This fits with out exploration earlier.

Our model is:
lm(price) = 0.415 + 9.144 x carat^(1/3) - 1.093 x carat + ... x cut + ... x
color + ... x clarity) + E

E being the error term


## Model Problems

Our data is from 2008. We need to account for inflation, and the diamond market
is quite different now than it was. Prices plummeted in 2008 due to the global
financial crisis. Since then prices, at least for wholesale polished diamonds,
have grown at about 6% per year, compound annual rate.

The rapidly growing number of couples in China buying diamond engagement rings
might also explain this increase.

Finally, diamonds prices grew unevenly across different carat sizes since 2008,
meaning the model we estimated couldn't simply be adjusted by inflation.